---
title: "Amusement Park"
output:
  pdf_document: default
    
  word_document: default
  html_document: default
---
# Introduction
This report extracts and analyse the data from [tidytuesday 2019-09-10 exercise](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-09-10)


# Texas Injuries
Load packages
```{r message=FALSE}
library(tidyverse)
library(lubridate)
```
Get the data from csv file
```{r message=FALSE}
tx_injuries <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/tx_injuries.csv")
```
Data examination
```{r}
head(tx_injuries)
```
Check relative injuries to different body_part
```{r}
head(unique(tx_injuries$body_part))
length(unique(tx_injuries$body_part))
```
Check rides
```{r}
head(unique(tx_injuries$ride_name))
length(unique(tx_injuries$ride_name))
```
I find that the columns body_part and ride_name do not have any generic discrete variables. It would be difficult to convert these columns into desired categorical variables.The column injuries by date seems to be an ideal candidate for analysis. In next steps, I select and rename required columns and plot injuries by date.
``` {r}
tx_injuries_selected <- tx_injuries %>%
     select(park = name_of_operation, 
            city, 
            state = st, 
            ride = ride_name, 
            body_part, 
            injury_type = alleged_injury, 
            date = injury_date) 
head(tx_injuries_selected)
```
Wrangle the date column
```{r}
dates_formatted = mdy(tx_injuries_selected$date)
dates_base_1899 = as.Date(as.numeric(tx_injuries_selected$date), origin = "1899-12-30")
date_wrangled = if_else(is.na(dates_formatted), dates_base_1899, dates_formatted)
tx_injuries_wrangled <- tx_injuries_selected %>% mutate(date = date_wrangled)
head(tx_injuries_wrangled)
```
Find injuries by date
```{r}
injuries_vs_date <- tx_injuries_wrangled %>% count(date)
head(injuries_vs_date)
```
Plot injuries by date
``` {r}
injuries_vs_date %>% ggplot(aes(x = date, y = n)) + 
      geom_line() + 
      theme_bw() + 
      labs(y = "Number of Injuries") + 
      guides(color = "none")
```
\newline

It can be observed that during the middle of every year (late spring and summer), there is a spike in the number of injuries.
# Safer parks
The data set is required to be tranformed into a tidy format. The package dplyr is required for sorting, filtering and summarizing data in this set.
```{r message=FALSE}
library("dplyr")
library("tidytext")
safer_parks <-
  readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-10/saferparks.csv")
head(safer_parks)
```
I need to restructure injury description as one token per row format, removing rows with stopwords. Finally I filter them into six possible body part buckets.
``` {r}
body_parts <- tolower(c("HEAD", "BACK", "EAR", "HIP", "ARM", "LEG"))
body_parts_freq <- safer_parks %>%
  unnest_tokens(output=word, input=injury_desc) %>% #
  anti_join(get_stopwords()) %>% # Remove stopwords
  filter(word %in% body_parts) %>% # Filter to rows with body part
  mutate(word = toupper(word)) %>% # Convert all string variables in word to upper case
  distinct(word, .keep_all = TRUE) %>% # Remove duplicate rows based on word
  mutate(word=as.factor(word)) %>% # Creating a new word datacolumn and categorizing the data into levels
  group_by(word) %>% # Converts word to a grouped table 
  summarise(total = sum(num_injured))
head(body_parts_freq)
```
It can be observed that head and back injuries are proportianately higher compared to other body part injuries.